---
title: "Relationship between Education and Years of Experience"
author: "Sophia Chiesa"
date: "`r Sys.Date()`"
output: pdf_document
---

## 1. Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggridges)
```

### Read in the data
```{r}
salary <- read_csv("Salary_Data.csv")
```

### Filter NAs
There are only 3 NA rows in the dataset so we will filter them out.
```{r}
salary %>% filter(is.na(`Salary`))

salary <- salary %>% filter(!is.na(`Salary`))
```

### Check on most common job titles
```{r}
salary %>% group_by(`Job Title`) %>%
  summarize(n_jobs = n()) %>%
  arrange(desc(n_jobs))
```

### Checking for salary outliers
```{r}
ggplot(data=salary) +
  geom_boxplot(mapping=aes(x=Salary))
```
There are no values of salary that are outliers (outside of 1.5 * IQR as shown in the boxplot).


## 2. Looking at the relationship between education level and salary
```{r}
salary %>% group_by(`Education Level`) %>%
  summarize(n_people = n())
```
Some data carpentry needs to be done to combine "Bachelor's" with "Bachelor's Degree", etc.

### Data carpentry to combine identical education levels
```{r}
salary <- salary %>% mutate(`Education Level` = case_when(
  `Education Level` == "Bachelor's Degree" ~ "Bachelor's",
  `Education Level` == "Master's Degree" ~ "Master's",
  `Education Level` == "phD" ~ "PhD",
  .default = `Education Level`)) %>%
  filter(!is.na(`Education Level`))

salary %>% group_by(`Education Level`) %>%
  summarize(n_people = n())
```

### Density plot of salary per education level
```{r}
# order education levels
salary$edu = factor(salary$`Education Level`, levels=c("High School", "Bachelor's", "Master's", "PhD"))

ggplot(data=salary) +
  geom_density_ridges(mapping=aes(x=`Salary`,
                                  y=edu,
                                  fill=edu),
                      alpha=0.5, show.legend=F) +
  labs(title="Salary distribution for each education level",
       y="Education Level",
       x="Salary (USD per year)") +
  scale_x_continuous(breaks=c(0, 50000, 100000, 150000, 200000, 250000))
```

### How does salary change with years of experience relative to education level?
```{r}
ggplot(data=salary) +
  geom_point(mapping=aes(y=`Salary`, x=`Years of Experience`),
             alpha=0.4) +
  facet_wrap(~`Education Level`) +
  labs(title="Salary vs. years of experience for each education level")
```
It is not clear from these scatterplots whether the interaction between education level and years of experience is a significant predictor of salary. However, it looks like the relationship between salary and YOE may not be exactly linear, especially for Master's and PhD groups. Therefore, I will look for a transformation for years of experience.


### Square root transformation of YOE
```{r}
salary <- salary %>% mutate(sqrt_yoe = sqrt(`Years of Experience`))

ggplot(data=salary) +
  geom_point(mapping=aes(y=`Salary`, x=sqrt_yoe),
             alpha=0.4) +
  facet_wrap(~`Education Level`) +
  labs(title="Salary vs. sqrt(years of experience) for each education level",
       x="Square root of years of experience")
```
To make the relationship more linear I have taken the square root of years of experience.


### Testing whether education level:years of experience is a significant predictor of salary
```{r}
lm.edu_yoe <- lm(`Salary` ~ sqrt_yoe * `Education Level`, data=salary)
summary(lm.edu_yoe)
```

### Plot the model with the 4 different lines
```{r}
bach <- salary %>% filter(`Education Level` == "Bachelor's")
high <- salary %>% filter(`Education Level` == "High School")
mast <- salary %>% filter(`Education Level` == "Master's")
phd <- salary %>% filter(`Education Level` == "PhD")

ggplot(bach, mapping=aes(x=sqrt_yoe, y=`Salary`)) +
  geom_point() +
  geom_abline(intercept=lm.edu_yoe$coefficients[1], slope=lm.edu_yoe$coefficients[2])

ggplot(high, mapping=aes(x=sqrt_yoe, y=`Salary`)) +
  geom_point() +
  geom_abline(intercept=lm.edu_yoe$coefficients[1] + lm.edu_yoe$coefficients[3],
              slope=lm.edu_yoe$coefficients[2] + lm.edu_yoe$coefficients[6])

ggplot(mast, mapping=aes(x=sqrt_yoe, y=`Salary`)) +
  geom_point() +
  geom_abline(intercept=lm.edu_yoe$coefficients[1] + lm.edu_yoe$coefficients[4],
              slope=lm.edu_yoe$coefficients[2] + lm.edu_yoe$coefficients[7])

ggplot(phd, mapping=aes(x=sqrt_yoe, y=`Salary`)) +
  geom_point() +
  geom_abline(intercept=lm.edu_yoe$coefficients[1] + lm.edu_yoe$coefficients[5],
              slope=lm.edu_yoe$coefficients[2] + lm.edu_yoe$coefficients[8])

plot(lm.edu_yoe, which=1)
plot(lm.edu_yoe, which=2)
```
The interaction is a significant predictor but looks like the residuals may not be normally distributed?


### Use an F-test to compare the above model to one without the interaction
```{r}
lm.edu_yoe2 <- lm(`Salary` ~ sqrt_yoe + `Education Level`, data=salary)
plot(lm.edu_yoe2, which=1)
plot(lm.edu_yoe2, which=2)

summary(lm.edu_yoe2)

anova(lm.edu_yoe, lm.edu_yoe2)
```

The F-test shows that at least one interaction term between education level and years of experience is significant. However, both the model with and without the interaction show the same issues of having residuals not centered around 0 and a non-normal qqplot.

## Maybe unrelated, but what are the typical education levels for selected job titles?
```{r}
p.edu_jobs <- salary %>%
  group_by(`Job Title`) %>%
  filter(`Job Title` == "Data Analyst" |
         `Job Title` == "Data Scientist" |
         `Job Title` == "Software Engineer") %>%
  summarize(count = n()) %>%
    ggplot(mapping=aes(x=`Job Title`, y=count, col=`Years of Experience`)) +
    geom_bar(position="stack", stat="identity")

p.edu_jobs
```

### 