---
title: "Relationship between Education and Years of Experience"
author: "Sophia Chiesa"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggridges)
```

```{r}
salary <- read_csv("Salary_Data.csv")
```

```{r}
salary %>% group_by(`Job Title`) %>%
  summarize(n_jobs = n()) %>%
  arrange(desc(n_jobs))
```


## Relationship between education level and salary
```{r}
salary %>% group_by(`Education Level`) %>%
  summarize(n_people = n())
```
There are only 3 NAs so I will filter these out. Some data carpentry needs to be done to combine "Bachelor's" with "Bachelor's Degree", etc.

### Data carpentry to combine identical education levels
```{r}
salary <- salary %>% mutate(`Education Level` = case_when(
  `Education Level` == "Bachelor's Degree" ~ "Bachelor's",
  `Education Level` == "Master's Degree" ~ "Master's",
  `Education Level` == "phD" ~ "PhD",
  .default = `Education Level`)) %>%
  filter(!is.na(`Education Level`))

salary %>% group_by(`Education Level`) %>%
  summarize(n_people = n())
```

### Density plot of salary per education level
```{r}
ggplot(data=salary) +
  geom_density_ridges(mapping=aes(x=`Salary`,
                                  y=`Education Level`,
                                  fill=`Education Level`),
                      alpha=0.5)
```

### How does salary change with years of experience relative to education level?
```{r}
p <- ggplot(data=salary) +
  geom_point(mapping=aes(y=`Salary`, x=`Years of Experience`),
             alpha=0.4)

p + facet_wrap(~`Education Level`)
```
It is not clear from these scatterplots whether the interaction between education level and years of experience is a significant predictor of salary.

### Testing whether education level:years of experience is a significant predictor of salary
```{r}
lm.edu_yoe <- lm(`Salary` ~ `Years of Experience` * `Education Level`, data=salary)
summary(lm.edu_yoe)
```

### Plot the model with the 4 different lines
```{r}
bach <- salary %>% filter(`Education Level` == "Bachelor's")
high <- salary %>% filter(`Education Level` == "High School")
mast <- salary %>% filter(`Education Level` == "Master's")
phd <- salary %>% filter(`Education Level` == "PhD")

ggplot(bach, mapping=aes(x=`Years of Experience`, y=`Salary`)) +
  geom_point() +
  geom_abline(intercept=lm.edu_yoe$coefficients[1], slope=lm.edu_yoe$coefficients[2])

ggplot(high, mapping=aes(x=`Years of Experience`, y=`Salary`)) +
  geom_point() +
  geom_abline(intercept=lm.edu_yoe$coefficients[1] + lm.edu_yoe$coefficients[3],
              slope=lm.edu_yoe$coefficients[2] + lm.edu_yoe$coefficients[6])

ggplot(mast, mapping=aes(x=`Years of Experience`, y=`Salary`)) +
  geom_point() +
  geom_abline(intercept=lm.edu_yoe$coefficients[1] + lm.edu_yoe$coefficients[4],
              slope=lm.edu_yoe$coefficients[2] + lm.edu_yoe$coefficients[7])

ggplot(phd, mapping=aes(x=`Years of Experience`, y=`Salary`)) +
  geom_point() +
  geom_abline(intercept=lm.edu_yoe$coefficients[1] + lm.edu_yoe$coefficients[5],
              slope=lm.edu_yoe$coefficients[2] + lm.edu_yoe$coefficients[8])

plot(lm.edu_yoe, which=1)
plot(lm.edu_yoe, which=2)
```
The interaction is a significant predictor but looks like the residuals may not be centered around 0 for high salary and that they may not be normally distributed?


### Use an F-test to compare the above model to one without the interaction
```{r}
lm.edu_yoe2 <- lm(`Salary` ~ `Years of Experience` + `Education Level`, data=salary)
plot(lm.edu_yoe2, which=1)
plot(lm.edu_yoe2, which=2)

summary(lm.edu_yoe2)

anova(lm.edu_yoe, lm.edu_yoe2)
```

The F-test shows that at least one interaction term between education level and years of experience is significant. However, both the model with and without the interaction show the same issues of having residuals not centered around 0 and a non-normal qqplot.

## Maybe unrelated, but what are the typical education levels for selected job titles?
```{r}
p.edu_jobs <- salary %>%
  group_by(`Job Title`) %>%
  filter(`Job Title` == "Data Analyst" |
         `Job Title` == "Data Scientist" |
         `Job Title` == "Software Engineer") %>%
  summarize(count = n()) %>%
    ggplot(mapping=aes(x=`Job Title`, y=count, col=`Years of Experience`)) +
    geom_bar(position="stack", stat="identity")

p.edu_jobs
```

