---
title: "STAT_423_project"
author: "Lorenzo, Henry, Sophia"
date: "2024-02-29"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(car)
library(tidyverse)
library(MASS)
Salary_Data <- read.csv("Salary_Data.csv")
Salary_Data <- na.omit(Salary_Data)
```

# Data description

The data used for this project is the Salary_Data dataset from Kaggle, found [here](https://www.kaggle.com/datasets/mohithsairamreddy/salary-data/data)

The description of the data set does not say where the data is from geographically. However, the data set creator did respond to a comment asking what unit the salary data was in, saying that it was in Indian Rupees(INR). Given this, it is most likely that this data set is from India.

There are 6698 complete rows and six columns: age (in years), gender (male, female, or other), education level (Bachelor’s, Master’s, or PhD), job title, years of experience, and salary in USD. The numeric variables are age, years of experience, and salary, while gender, education level, and job title are categorical. While education level and gender have predetermined levels, the job title column is freeform; some job titles come up hundreds of times while others only appear once. Additionally, there are only 14 rows with gender as other, so there is relatively less data for that point.

Due to the job title column having a large number of categories, we will be grouping categories together to have a more reasonable number of categorical levels for our analysis. Given the arbitrary nature of this, it could potentially have an impact on the significance of our findings of if job title is a significant predictor of salary.

# Job Categorization

For our analysis of the Salary Dataframe, we wanted to do linear regression of the salary based on the type of job someone was employed in. The first instinct would be to use the `Job.Title` column of our dataframe. The problem with this approach is that there are way too many job titles that were listed in the data there were `r length(Salary_Data %>% count(Job.Title)$Job.Title)` different job titles listed in the dataset, which would make any categorical regression nigh impossible.

So we decided to categorize our job titles into different job categories which correspond to different kinds of jobs. The 5 levels for our `Job.Category` variable are `STEM, BUSINESS, ADMIN, SERVICE, and OTHER`. 

- `STEM` describes scientific and technical jobs, like computer programming, scientific research, and engineering. 

- `BUSINESS` refers to any jobs involving the financial and marketing aspects of running businesses and companies. 

- `ADMIN` refers to managerial and administrative jobs, involves project management, training, and human resource management. 

- `SERVICE` refers to jobs the involve mainly interacting with customers and helping them in some way

- `OTHER` refers to jobs that don't fit any of the categories listed above

```{r}
# Job Category Word Lists
# STEM
STEM <- c(" Technology ", " Digital ", " Science ", " Engineer ", " Software ",
          " Network ", " Architect ", " Scientist ", " Engineering ", " Research ",
          " Data ", " Specialist ", " IT ", " Web ", " Developer ",
          " Stack ", " Researcher ", " Techinical ")
# Admin
ADMIN <- c(" Admin ", " Rep ", " Human ", " Recruiter ", " Relations ",
           " Strategy ", " Manager ", " Director ", " Management ", " Coordinator ",
           " Coordinator ", " Office ", " Training ", " Supply ", " VP ",
           " Accountant ", " Resources ", " Principal ", " Executive ", " Administrative ",
           " Officer ", " Entry ", " HR ", " Operations ", " Project ")

# Business
BUSINESS <- c(" Account ", " Business ", " Product ", " Capital ", " UX ", " Advertising ", " Quality ", " Development ", " Marketing ", " Finance ", " Sales ", " Financial ", " Associate ", " Social ", " Intelligence ", " Copywriter ", " Advisor ", " Media ", " Public ")

# Service
SERVICE <- c(" Customer ", " Success ", " Delivery ", " Service ", " Help ", " Support ")

# Function to Test if there is Job Category Overlap for Job Titles
is_STEM <- function(title) {
  title_standard <- paste("", title, "")
  for (word in STEM) {
    if(grepl(word, title_standard)) {
      return(TRUE)
    }
  }
  return(FALSE)
}

is_ADMIN <- function(title) {
  title_standard <- paste("", title, "")
  for (word in ADMIN) {
    if(grepl(word, title_standard)) {
      return(TRUE)
    }
  }
  return(FALSE)
}

is_BUSINESS <- function(title) {
  title_standard <- paste("", title, "")
  for (word in BUSINESS) {
    if(grepl(word, title_standard)) {
      return(TRUE)
    }
  }
  return(FALSE)
}

is_SERVICE <- function(title) {
  title_standard <- paste("", title, "")
  for (word in SERVICE) {
    if(grepl(word, title_standard)) {
      return(TRUE)
    }
  }
  return(FALSE)
}

check_overlap <- function(title) {
  bool_vector <- c()
  bool_vector[1] <- is_STEM(title)
  bool_vector[2] <- is_ADMIN(title)
  bool_vector[3] <- is_BUSINESS(title)
  bool_vector[4] <- is_SERVICE(title)
  if(mean(bool_vector) > 1/4) {
    return(TRUE)
  }
  return(FALSE)
}

# Takes care of overlapping categorizations for job titles
STEM.overlap <- c("Chief Data Officer", "Chief Technology Officer", "Data Entry Clerk", "Director of Data Science", "Director of Engineering", "IT Manager", "IT Support", "IT Support Specialist", "Principal Engineer", "Principal Scientist", "Research Director", "Senior IT Project Manager", "Senior IT Support Specialist", "Senior Project Engineer", "Software Engineer Manager", "Software Manager", "Software Project Manager", "Technical Support Specialist", "UX Researcher")
ADMIN.overlap <- c("Director of Human Capital", "Project Engineer", "Senior HR Specialist", "Senior Human Resources Specialist", "Senior Training Specialist", "Training Specialist")
BUSINESS.overlap <- c("Account Manager", "Business Development Manager", "Content Marketing Manager", "Digital Marketing Manager", "Digital Marketing Specialist", "Director of Business Development", "Director of Finance", "Director of Marketing", "Director of Product Management", "Director of Sales", "Director of Sales and Marketing", "Financial Manager", "Junior Account Manager", "Junior Advertising Coordinator", "Junior Business Operations Analyst", "Junior Marketing Coordinator", "Junior Marketing Manager", "Junior Marketing Specialist", "Junior Product Manager", "Junior Social Media Manager", "Junior Social Media Specialist", "Marketing Coordinator", "Marketing Director", "Marketing Manager", "Marketing Specialist", "Product Manager", "Product Marketing Manager", "Public Relations Manager", "Sales Director", "Sales Executive", "Sales Manager", "Sales Operations Manager", "Senior Account Executive", "Senior Account Manager", "Senior Business Development Manager", "Senior Financial Manager", "Senior Marketing Coordinator", "Senior Marketing Director","Senior Marketing Manager", "Senior Marketing Specialist", "Senior Product Development Manager", "Senior Product Manager", "Senior Product Marketing Manager", "Senior Sales Manager", "Social Media Manager", "Social Media Specialist", "VP of Finance")
SERVICE.overlap <- c("Customer Service Manager", "Customer Service Rep", "Customer Success Manager", "Customer Success Rep", "Junior Customer Support Specialist")
sort.overlap <- function(job_titles) {
  i <- 1
  categories <- c()
  for(title in job_titles) {
    if(title %in% STEM.overlap) {
      categories[i] <- "STEM"
    } else if(title %in% ADMIN.overlap) {
      categories[i] <- "ADMIN"
    } else if(title %in% BUSINESS.overlap) {
      categories[i] <- "BUSINESS"
    } else if(title %in% SERVICE.overlap) {
      categories[i] <- "SERVICE"
    } else {
      categories[i] <- title
    }
    i <- i + 1
  }
  return(categories)
}

job_category <- function(job_titles) {
    i <- 1
    categories <- c()
    for (title in job_titles) {
      if(title %in% STEM.overlap) {
        categories[i] <- "STEM"
      } else if(title %in% ADMIN.overlap) {
        categories[i] <- "ADMIN"
      } else if(title %in% BUSINESS.overlap) {
        categories[i] <- "BUSINESS"
      } else if(title %in% SERVICE.overlap) {
        categories[i] <- "SERVICE"
      } else {
        categories[i] <- case_when(check_overlap(title) ~ "OVERLAP",
                        is_STEM(title) ~ "STEM",
                        is_ADMIN(title) ~ "ADMIN",
                        is_BUSINESS(title) ~ "BUSINESS",
                        is_SERVICE(title) ~ "SERVICE",
                        TRUE ~ "OTHER")
      }
      i <- i + 1
    }
    return(categories)
}
```

```{r}
job_categories <- job_category(Salary_Data$Job.Title)
# job_categories <- sort.overlap(job_categories)
Salary_Data <- Salary_Data %>% mutate(Job.Category = job_categories)
```

# Education

We also noticed later on that the Education.Level column had categories that should be put together, such as phD and PhD, and Bachelor's and Bachelor's Degree, where they denoted the same level of education. So, we fixed that to the levels of High School, Bachelor's, Master's, and PhD.

```{r}
#Ed.LVLs <- unique(Salary_Data$Education.Level)
#Ed.LVLs

High <- c("High School")
Bachelor <- c("Bachelor's", "Bachelor's Degree")
Master <- c("Master's", "Master's Degree")
PhD <- c("PhD", "phD")


education <- function(ed_lvls) {
    i <- 1
    ED <- c()
    n <- length(ed_lvls)
    while (i <= n) {
      if(ed_lvls[i] %in% High){
        ED[i] <- "High School"
      } else if(ed_lvls[i] %in% Bachelor){
        ED[i] <- "Bachelor's"
      } else if(ed_lvls[i] %in% Master) {
        ED[i] <- "Master's"
      } else if(ed_lvls[i] %in% PhD) {
        ED[i] <- "PhD"
      }else {
        ED[i] <- NA
      }
      i <- i + 1
    }
    return(ED)
}
```

```{r}
schooling <- education(Salary_Data$Education.Level)
Salary_Data$Education <- schooling
Salary_Data <- na.omit(Salary_Data)
```


# Service


# starting model


# collinearity of age and experience

We want to check if age and years of experience are collinear. It makes logical sense that they would be, as you would not expect a person fresh out of college to have much experience. And a person who is older could have been is a job for years. But, there could be other factors that cause them to not be related, such as changing jobs or fields, and thus being older with less experience.

We will use three methods to check the collinearity of age and experience. The first will be a scatterplot of the two variables. If they are unrelated we will see an even spread, and if they are related, there will be a more linear spread. We will also use two functions in R, to find two numbers that would hint towards collinearity, the correlation coefficient and the variance inflation factor. If the correlation coefficient is greater than 0.8 or the variance inflation factor is greater than 10, there is evidence of collinearity. The third test will be fitting a linear regression to just age and experience, with age as the predictor and experience as the response. If the two are collinear, then there will be statistically significant evidence of a linear relationship between the two variables from a t-test.

```{r}
AgeVsExp <- Salary_Data[,c("Salary", "Age", "Years.of.Experience")]
AgeVsExp$Age <- as.numeric(AgeVsExp$Age)
AgeVsExp$Years.of.Experience <- as.numeric(AgeVsExp$Years.of.Experience)
Age_S <- as.numeric(AgeVsExp$Age)
Experience_S <- as.numeric(AgeVsExp$Years.of.Experience)

plot(Age_S, Experience_S, main = "Scatterplot of Age vs Experience",
     xlab = "Age",
     ylab = "Experience")
```
Looking at the scatterplot of age versus experience, we see that there appears to be a linear relationship between the two.

```{r echo=TRUE}
cor(AgeVsExp$Age, AgeVsExp$Years.of.Experience, use = "complete.obs")
```

Using the correlation function in R gives us a correlation value of 0.9377, which is greater than what we said was needed to consider a collinear relationship possible.

```{r echo=TRUE}
AVE.lm <- lm(Salary ~ Age + Years.of.Experience, data = AgeVsExp)
vif(AVE.lm)
```
Creating a quick linear model of Salary predicted by Age and Experience, we can find the Variance Inflation Factor of the two predictors. We see that each of them have a VIF of greater than 8. This again shows us that there is a strong case for collinearity between the two factors.


```{r}
AVE.lm2 <- lm(Years.of.Experience ~ Age, data = AgeVsExp)
summary(AVE.lm2)
```
Using R to create a linear regression between Age and Years of Experience we see that the t-test gives a very low p-value, less than 0.001. This means that there is strong evidence of a linear relationship between the two variables, matching all of our previous tests. We see that if Age increases by one, years of experience will increase by roughly 0.746. Thus, as we have statistically significant evidence of Age and Years of Experience being collinear, we will not have both of them in our final model.

# Checking for interactions

To check for interactions, we fit a linear model including all predictors, and all of their interactions. From those, we selected only the interactions that were statistically significant at the 0.05 level of probability. This gives us the following interactions as significant. Education has multiple significant interactions with both age and experience. Education also has multiple significant interactions with job category. And, looking at rows 16 to 19, gender appears to have a significant interaction with education, and years of experience.

```{r}
inter <- lm(Salary ~ (Age + Gender + Education + Years.of.Experience + Job.Category)^2, data = Salary_Data)
inter_summary <- summary(inter)

# Get the coefficient matrix
coefs <- inter_summary$coefficients

# Identify the significant variables
vars <- rownames(coefs)[which(coefs[, 4] < 0.05)]
#vars

significant <- data.frame(matrix(NA, nrow = 30, ncol = 2))
colnames(significant) <- c("predictor", "p value")
significant$predictor <- vars

r <- 1
for(i in 1:57){
  if(coefs[i,4] < 0.05) {
    significant[r,2] <- coefs[i,4]
    r <- r + 1
  }
}

tail(significant, n = 21)
```

So, to check that these interactions are significant, we will use the AIC and BIC functions. We will start with a fit of salary based on gender, experience, job category and education. I am leaving out age, as we know it to be collinear with experience based on work done above. We will compare this fit to one that also includes the interactions of education with experience, job category, and gender individually; it will also include the interaction between gender and experience. If the AIC and BIC for the fit including interactions is lower, we will conclude that the interactions are significant and should be included in our final model. Looking at our results below, this does hold, so we will add the interactions to our final model.

```{r}
fit_basic <- lm(Salary ~ Gender + Years.of.Experience + Job.Category + Education, data = Salary_Data)
fit_interactions <- lm(Salary ~ Gender + Years.of.Experience + Job.Category + Education + Education:Years.of.Experience + Education:Job.Category + Education:Gender + Gender:Years.of.Experience, data = Salary_Data)

AIC(fit_basic, fit_interactions)
BIC(fit_basic, fit_interactions)
```



# Education level impact


# Outliers


# Final model